---
title: "titanic"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Exploring Titanic Dataset 
Look at working directory and read the data

```{r titanic}
rm(list=ls()) 
train <- read.csv("train.csv",stringsAsFactors = FALSE) 
str(train)
summary(train)
```
## How many people survived

```{r}
survived <- table(train$Survived)
names(survived) <- c('No', 'Yes')
survived
```

## get proportion table

```{r pressure, echo=FALSE}
prop.table(survived)
```

## enter dying prediction to testing dataset
```{r}
test <- read.csv("test.csv",stringsAsFactors = FALSE) 
test$Survived <- rep(0, 418)
```

## write first submission 

```{r}
submit <- data.frame(PassengerId = test$PassengerId, Survived = test$Survived)
write.csv(submit, file = "theyallperish.csv", row.names = FALSE)
```
## We get sex to see who survived
```{r}
prop.table(table(train$Sex,train$Survived),1)
test$Survived <- 0
#install.packages("dplyr")
#library(dplyr)
#test = select(test, -s)
```
## having that females survived with a higher percentage I can say first females survived
```{r}
test$Survived[test$Sex=='female']<-1

writemycsv <- function (myData){
  submit <- data.frame(PassengerId = myData$PassengerId, Survived = myData$Survived)
  write.csv(submit, file = "theyallperish.csv", row.names = FALSE)
}
writemycsv(test)
```
## I start contemplating age value into the picture and create children feature
```{r}
summary(train$Age)
train$child <- 0
train$child[train$Age<18] <- 1
childFix <- function(tempData){
  tempData$child <- 0
  tempData$child[tempData$Age<18] <- 1
  return(tempData)
}
```
## we can group survived passenger that are child by sex
we can see most female childs survived
```{r}
aggregate(Survived~child+Sex,data=train,FUN = sum)
aggregate(Survived~child+Sex,data=train,FUN = length)
aggregate(Survived~child+Sex,data=train,FUN =function(x){sum(x)/length(x)})
```
## create fare feature to see it will help with output
```{r}
summary(train$Fare)
fareFix <- function(tempData){
  tempData$fareCat <- '<10'
  tempData$fareCat[tempData$Fare<=10 & tempData$Fare<=20] <- '10-20'
  tempData$fareCat[tempData$Fare>20 & tempData$Fare<=30] <- '21-30'
  tempData$fareCat[tempData$Fare>31] <- '>31'  
  return(tempData)
}
train$fareCat <- '<10'
train$fareCat[train$Fare<=10 & train$Fare<=20] <- '10-20'
train$fareCat[train$Fare>20 & train$Fare<=30] <- '21-30'
train$fareCat[train$Fare>31] <- '>31'
aggregate(Survived~fareCat+Pclass+Sex,data=train,FUN=function(x){sum(x)/length(x)})
```
## having a viewed that class 3 females did not do good we do a new prediction
```{r}
test$Survived <- 0
test$Survived[test$Sex=='female'] <-1
test$Survived[test$Sex=='female'&test$Pclass==3&test$Fare>=20] <-0
writemycsv(test)
```
## decision trees are going to be used instead of picking and fishing for results
```{r}
library(rpart)
#method class we get a 0 or 1, method anova returns decimal quantaties
fit <- rpart(Survived ~ Pclass + Sex + Age + SibSp + Parch + Fare + Embarked, data=train,method = "class")
#plot did not work as expected 
plot(fit)
survived3 <- predict(fit,test, type = "class")
test <- data.frame(PassengerId = test$PassengerId, Survived=survived3)
writemycsv(test)
```
## will try feature engineering starting with the name
#rm(list=ls())  
```{r}
train$Name[1]
train <- read.csv("train.csv")
test <- read.csv("test.csv")
#test <- read.csv("test.csv",stringsAsFactors = FALSE) 
test$Survived <- NA
#add child column to test
train <- childFix(train)
test <- childFix(test)
#add fare column to test
train <- fareFix(train)
test <- fareFix(test)

combi <- rbind(train,test)
combi$Name <- as.character(combi$Name)
combi$Name[1]
#we obtain part of name we want
strsplit(combi$Name[1],split = '[,.]')[[1]][2]
#create a feature just containing the tittle of the preson
combi$Title <- sapply(combi$Name, FUN = function(x){strsplit(x,split = '[,.]')[[1]][2]})
#clean title before moving on
combi$Title <- sub(' ','',combi$Title)
table(combi$Title)
combi$Title[combi$Title %in% c('Mlle','Mme')] <- 'Mlle'
combi$Title[combi$Title %in% c('Capt','Don','Major','Sir','Jonkheer')] <- 'Sir'
combi$Title[combi$Title %in% c('Dona','Lady','the Countess')] <- 'Lady'
combi$Title <- factor(combi$Title)
combi$FamilySize <- combi$Parch + combi$SibSp + 1
combi$Surname <- sapply(combi$Name, FUN=function(x) {strsplit(x, split='[,.]')[[1]][1]})
combi$FamilyID <- paste(as.character(combi$FamilySize), combi$Surname, sep="")
combi$FamilyID[combi$FamilySize <= 2] <- 'Small'
table(combi$FamilyID)
famIDs <- data.frame(table(combi$FamilyID))
famIDs <- famIDs[famIDs$Freq <= 2,]
combi$FamilyID[combi$FamilyID %in% famIDs$Var1] <- 'Small'
combi$FamilyID <- factor(combi$FamilyID)
train <- combi[1:891,]
test <- combi[892:1309,]
```
## after creating new features and slicing combi back into the original datasets a new model is created
```{r}
fit2 <- rpart(Survived ~ Pclass + Sex + Age + SibSp + Parch + Fare + Embarked + Title + FamilySize + FamilyID, data=train,method = "class")
plot(fit2)
survived4 <- predict(fit2,test, type = "class")
test2 <- data.frame(PassengerId = test$PassengerId, Survived=survived4)
writemycsv(test2)
```
## we go back and check age
there are 263 values that are NA
```{r}
summary(combi$Age)
Agefit <- rpart(Age ~ Pclass + Sex + SibSp + Parch + Fare + Embarked + Title,
                  data=combi[!is.na(combi$Age),],
                  method="anova")
combi$Age[is.na(combi$Age)] <- predict(Agefit, combi[is.na(combi$Age),])
summary(combi)
summary(combi$Embarked)
train <- combi[1:891,]
test <- combi[892:1309,]
fit3 <- rpart(Survived ~ Pclass + Sex + Age + SibSp + Parch + Fare + Embarked + Title + FamilySize + FamilyID, data=train,method = "class")
plot(fit3)
survived5 <- predict(fit3,test, type = "class")
test2 <- data.frame(PassengerId = test$PassengerId, Survived=survived5)
writemycsv(test2)
##put an embarqued place on everybody
summary(combi$Embarked)
which(combi$Embarked == '')
combi$Embarked[c(62,830)] = "S"
combi$Embarked <- factor(combi$Embarked)
##clear the na in fare
summary(combi$Fare)
which(is.na(combi$Fare))
combi$Fare[1044] <- median(combi$Fare, na.rm=TRUE)
```
##
```{r}
combi$FamilyID2 <- combi$FamilyID
combi$FamilyID2 <- as.character(combi$FamilyID2)
summary(combi$FamilySize)
combi$FamilyID2[combi$FamilySize <= 3] <- 'Small'
combi$FamilyID2 <- factor(combi$FamilyID2)
install.packages('randomForest')
library(randomForest)
set.seed(415)
train <- combi[1:891,]
test <- combi[892:1309,]
fit <- randomForest(as.factor(Survived) ~ Pclass + Sex + Age + SibSp + Parch + Fare +
                                            Embarked + Title + FamilySize + FamilyID2,
                      data=train, 
                      importance=TRUE, 
                      ntree=2000)
varImpPlot(fit)
```
## we check prediction with what we have so far
```{r}
 Prediction <- predict(fit, test)
submit <- data.frame(PassengerId = test$PassengerId, Survived = Prediction)
writemycsv(submit)
install.packages('party')
library(party)
set.seed(415)
fit <- cforest(as.factor(Survived) ~ Pclass + Sex + Age + SibSp + Parch + Fare +
                                       Embarked + Title + FamilySize + FamilyID,
                 data = train, 
                 controls=cforest_unbiased(ntree=2000, mtry=3))
Prediction <- predict(fit, test, OOB=TRUE, type = "response")
submit <- data.frame(PassengerId = test$PassengerId, Survived = Prediction)
writemycsv(submit)
summary(train$Sex)
```

## end approach finishes with a score of 0.81339
```{r}
library(dplyr)
glimpse(train)
```

